Usually, public health organizations considers {\color{red}(consider?)} vaccination as
{\color{red}(a?)} principal
preventive action again infectious disease, but is plausible only when
represents an economic reward. 
Due to the limited resources associated with vaccination programs, is imperative
to optimize the use of available resources. Using optimal control theory, we
formulate a vaccination schedule. The goal is to minimize the number of infected
persons and the cost of vaccine during a fixed time, to this example, we 
optimize the functional
$$
  J(u):= 
  \int_{0}^{T}
    A I(t) + u^2(t) .
$$
Here $u$ is the vaccination control and denotes the fraction of
susceptible individuals to vaccinate per unit of time. Since manage infected 
population imply resource consume, $A$ means the regarding cost per individual.
We also need a spreads dynamics. So, let $S(t)$, $E(t)$, $I(t)$, $R(t)$
respectively be the number of susceptible, exposed, infectious, and recovered
(immune) individuals at time $t$. Since vaccination of
the entire susceptible population is impractical, the model considers 
$0 \leq u(t) \leq 0.9$. Then the whole population $N$ is given by 
$N(t) = S(t) + E(t) + I(t) + R (t)$, and obeys
$
  \dot{N}(t) =
    (b - d)N(t) - aI(t).
$
Since $b$ is the recruitment rate and $d$ natural death, the term $b-d$ denotes 
the growth of the entire population. Then, the optimal control problem reads
%
%
\begin{equation} \label{eqn:epidemics_lenhart}
  \begin{aligned}
    \min_{u} & \int_{0}^{T} AI(t) + u^{2}(t) dt,
    \\
    \text{subject to}
    \\
      \dot{S}(t) &=
          bN(t) - dS(t) - cS(t)I(t) - u(t)S(t), \quad S(0) = S_0 \geq 0,   \\
      \dot{E}(t) &=
          cS(t)I(t) - (e + d)E(t), \quad E(0) = E_0 \geq 0,    \\
      \dot{I}(t) &=
          eE(t) - (g + a +d)I(t), \quad I(0) = I_0 \geq 0,     \\
      \dot{R}(t) &=
          gI(t) -dR(t) + u(t)S(t), \quad R(0) = R_0 \geq 0,    \\
        \dot{N}(t) &=
          (b - d)N(t) - aI(t), \quad N(0)= S_0 + E_0 + I_0 + R_0, 
  \end{aligned}
\end{equation}
see \Cref{tbl:epidemics_lenhart_des} for a parameters description.

\begin{table}[H]
  \begin{center}
    \begin{tabular}{rl}
      \toprule
        & \multicolumn{1}{c}{\textbf{Description}} 
        \\
      \midrule
        $b$
          & Recruitment rate
        \\
        $a$, $d$ 
          & Disease and natural death rates
        \\
        $c$
          & Incidence of disease
        \\
        $e$
          & Rate at which the exposed 
          \\
          & individuals become infectious
        \\
        $g$
          & Recovering rate
        \\
        $A$
          & Vaccination cost
        \\
        $T$
          & Final time
        \\
      \bottomrule
    \end{tabular}
    \caption{Parameters and simulation values of the epidemic model
      \eqref{eqn:epidemics_lenhart}.}
    \label{tbl:epidemics_lenhart_des}
  \end{center}
\end{table}
%