\subsection{Optimal Control Software}
  In addition to the implementation of the schemes discussed above, here we
  provide a list with useful software and some of its references. Here we follow
  the list reported by Rodrigues, Monteiro, and Torres in the preprint
  \cite{Rodrigues2014a}, see this reference for code examples and more details.
\medskip
\paragraph{OC-ODE}
  The \verb|OC-ODE| \cite{Gerdts2009}, Optimal Control of Ordinary-Differential 
Equations, by Matthias Gerdts, is a collection of Fortran 77 routines for 
optimal control problems subject to ordinary differential equations. It uses an 
automatic direct discretization method for the transformation of the OC problem 
into a finite-dimensional non linear problem, OC-ODE includes procedures for 
numerical adjoint estimation and sensitivity analysis. 
\medskip
\paragraph{DOTcvp}
  \citet{Hirmajer2009}, provide the MATLAB Toolbox \verb|DOTcvp|. Giving a 
piecewise solution for the control, the toolbox uses the control vector 
parametrization approach for the calculation of the  optimal control profiles. 
\medskip
\paragraph{Muscod-II}
  
  \verb|MUSCOD-II| is a robust and efficient optimization tool that allows to
quickly implement and solve very general optimal control problems in 
differential-algebraic equations (DAE). This package relies on the Multiple 
Shooting method for the solution of mixed integer nonlinear ODE or DAE. The 
authors%: \citeauthor*{Diehl2001}, 
provide the code and a reference manual 
\cite{Diehl2001}.
\medskip
\paragraph{Ipopt}
  \citet*{Wachter2006} provide the software package Ipopt (Interior Point 
  OPTimizer). Ipopt implements a primal-dual interior point method and uses a 
  line search strategy based on filter method and is written in \verb|Fortran| 
  and \verb|C|.
\medskip  
\paragraph{Knitro}
    Byrd, Nocedal, and Waltz r reports in \cite{Byrd2006} \verb|Knitro 5.0|, 
a C-package for nonlinear optimization that combines complementary approaches 
to nonlinear optimization to achieve robust performance over a wide range of 
application requirements. The package  is designed for solving large-scale, 
smooth nonlinear programming problems, and it is also effective for the 
following special cases: unconstrained optimization, nonlinear systems of 
equations, least squares, and linear and quadratic programming. Various 
algorithmic options are available, including two interior methods and an 
active-set method.  